Compact linkable ring signatures and applications

### 2019-09-24 paper/mrl.tex
Updates to CLSAG paper

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{assu}{Assumption}[section]

\title{Compact linkable ring signatures and applications}
\authors{Brandon Goodell\footnote{\texttt{surae.noether@protonmail.com}} and Sarang Noether\footnote{\texttt{sarang.noether@protonmail.com}} and \texttt{RandomRun}}
\affiliations{Monero Research Lab}
\date{\today}
\type{RESEARCH BULLETIN}
\ident{MRL-0011}

@@

\begin{abstract}
We describe an efficient linkable ring signature scheme, compact linkable spontaneous anonymous group (CLSAG) signatures, for use in confidential transactions. Compared to the existing signature scheme used in Monero, CLSAG signatures are both smaller and more efficient to generate and verify for ring sizes of interest. We generalize the construction and show how it can be used to produce signatures with coins of different type in the same transaction.
\end{abstract}

\section{Introduction}
Ring signatures are used in Monero to provide transaction signer ambiguity. Originally, a simple Schnorr-like linkable ring signature \cite{liu2004linkable} was used that included previous one-time output public keys as possible signers. Later, the addition of Pedersen commitments to hide amounts necessitated a generalization to the signature \cite{noether2016ring} in order to prove transaction balance without revealing the signer.

The current ring signature construction used in Monero, MLSAG signatures, scale linearly with the number of ring members included in the signature. In particular, they require the inclusion of two scalars per ring member.

We note that the MLSAG signature scheme, while general enough to permit linking across multiple sets of input keys, is in practice used as a hybrid approach to a linkable ring signature. Indeed, each signature input consists of two keys: the first is a standard one-time output public key, and the second is a key computed using the associated Pedersen commitments. Linkability is only used for the first key, to ensure that double-spending does not occur.

This construction is flexible. For example, the format can be used for usual ring confidential transactions or could be modified to describe \textit{colored ring confidential transactions} with multiple so-called coin colors, allowing users to transact with multiple assets simultaneously.

\subsection{Our contribution}

In this work, we introduce a $d$-dimensional linkable ring multisignature ($d$-LRMS) scheme suitable for use in a colored ring confidential transaction scheme with $d$ colors. Our scheme is \textit{compact} in the sense that signature size scales with the sum of ring size and the dimension $d$; that is, increasing $d$ yields the same additional number of signature elements, regardless of the ring size. We prove our scheme is unforgeable up to the hardness of the $k$-one-more discrete logarithm problem.

We call our signature scheme $d$-dimensional compact linkable spontaneous anonymous group ($d$-CLSAG) signatures. We demonstrate how to use this scheme for ring confidential transactions in Monero by replacing multi-layered linkable spontaneous anonymous group (MLSAG) signatures with $d$-CLSAG signatures.  The resulting scheme is more efficient in both signature size and verification time than current Monero transaction structures, and can be seamlessly integrated into the Monero transaction protocol.

\section{Notation}

In this document, we denote algorithms with typefont majuscule English letters like $\texttt{A}$, $\texttt{B}$, or $\texttt{O}$, or typefont names like $\texttt{Setup}$, $\texttt{KeyGen}$, and so on.

Group parameters are denoted as a tuple $(p, \mathbb{G}, d, G)$ where $\mathbb{G}$ is an elliptic curve group with prime order $p$, $d$ is a dimension, and $G$ is a generator of $\mathbb{G}$. We denote integers, bits, indices, and scalars in $\mathbb{Z}/p\mathbb{Z}$ with minuscule English letters $x, y, z, b, c, i, j, k$, etc.\ and we denote group elements with majuscule English letters, $G, X, W$, and so on. We use miniscule Greek letters like $\sigma$ to describe signatures and majuscule calligraphic Latin letters like $\mathfrak{T}$ when describing linkability tags.

@@

We distinguish oracles with calligraphic font, e.g.\ $\mathcal{CO}$ denotes a corruption oracle, $\mathcal{SO}$ denotes a signing oracle. If the codomain of a random oracle is the field of scalars $\mathbb{Z}/p\mathbb{Z}$, we denote this $\mathcal{H}^s$ (hash-to-scalar). If the codomain is $\mathbb{G}$, we denote this $\mathcal{H}^p$ (hash-to-point).

\section{Linkable ring multisignatures}

In this section, we recall linkable ring signature (LRS) schemes, we modify the definition to account for $d$-dimensional keys for linkable ring multisignature (LRMS) schemes, and we provide two examples. We emphasize that our description does not present multisignatures as computed between mutually untrusted parties. We use the term to mean a $d$-of-$d$ threshold multisignature with all keys controlled by the same signer; in particular, we have none of the communication steps necessary for computing a multisignature.

\subsection{LRS and LRMS schemes}

\begin{defn}[LRS]\label{def:lrs}
A \textit{linkable ring signature scheme} is a tuple $(\texttt{Setup}, \texttt{KeyGen}, \texttt{Sign}, \texttt{Verify}, \texttt{Link})$ satisfying the following. 
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ takes as input a security parameter $1^\lambda$, produces some public parameters $\textit{par}$.

@@

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, sk)\right) \to \left\{\bot_{\texttt{Sign}}, (\sigma, \mathfrak{T})\right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, an arbitrary message $m \in \left\{0,1\right\}^*$, an ad hoc \textit{ring} of public keys $\textbf{pk} = \left\{pk_1, \ldots, pk_n\right\}$, and a secret key $sk$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\textit{out} = \bot_{\texttt{Sign}}$ or a signature-tag pair $\textit{out} = (\sigma, \mathfrak{T})$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, (\sigma, \mathfrak{T}))\right) \to \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a ring of public keys $\textbf{pk}$, and a signature-tag pair $(\sigma, \mathfrak{T})$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$.

\item $\texttt{Link}\left(1^\lambda, \textit{par}, (\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)\right)$. $\texttt{Link}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, and a pair of signature-tag pairs $(\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)$. $\texttt{Link}$ produces as output a bit $b \in \left\{0,1\right\}$. 
\end{itemize}
\end{defn}

We extend this to a scheme with $d$-dimensional keys.

\begin{defn}[$d$-LRMS]
A \textit{$d$-dimensional linkable ring multisignature scheme} is a tuple of algorithms $(\texttt{Setup}, \texttt{KeyGen}, \texttt{Sign}, \texttt{Verify}, \texttt{Link})$ satisfying the following. 
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ works as described in \ref{def:lrs}.

\item $\texttt{KeyGen}(1^\lambda, \textit{par}) \to (\textbf{sk}, \textbf{pk})$. $\texttt{KeyGen}$ takes as input a security parameter $1^\lambda$ and public parameters $\textit{par}$. $\texttt{KeyGen}$ produces as output a private-public keypair $(\textbf{sk}, \textbf{pk})$ where each key is a $d$-dimensional vector. We refer to the first coordinate of each of these keys as the \textit{linking key} and the $d-1$ remaining coordinates (if there are any) are the \textit{auxiliary keys}.

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \underline{\textbf{pk}}, \textbf{sk})\right) \to \left\{\bot_{\texttt{Sign}}, (\sigma, \mathfrak{T})\right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a vector of public keys $\underline{\textbf{pk}} = \left\{\textbf{pk}_1, \ldots, \textbf{pk}_n\right\}$ where each $\textbf{pk}_i$ is a $d$-dimensional public key, and a $d$-dimensional secret key $\textbf{sk}$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature-tag pair $(\sigma, \mathfrak{T})$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \underline{\textbf{pk}}, (\sigma, \mathfrak{T}))\right) \rightarrow \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a matrix of public keys $\underline{\textbf{pk}}$, and a signature-tag pair $(\sigma, \mathfrak{T})$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$.

\item  $\texttt{Link}\left(1^\lambda, \textit{par}, (\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)\right)$. $\texttt{Link}$ works as described in \ref{def:lrs}.
\end{itemize}
\end{defn}

We refer to the linkability tag as the \textit{key image} for consistency with terminology in Monero. We also use the following terminology if the scheme satisfies the respective properties.
\begin{description}
\item[Correctly verified:] For any message $m$, $d$-dimensional secret key $\textbf{sk}$ with corresponding public key $\textbf{pk}$, and appropriately-sized matrix $\underline{\textbf{pk}} = \left\{\textbf{pk}_1, \ldots, \textbf{pk}_n\right\}$, if there exists an index $1 \leq \ell \leq n$ such that $\textbf{pk} = \textbf{pk}_\ell$, then $\texttt{Verify}(\texttt{Sign}(m, \underline{\textbf{pk}}, \textbf{sk})) = 1$.

\item[Correctly linkable:] For any messages $m$, $m^\prime$, any $d$-dimensional secret keys  $\textbf{sk}$ and $\textbf{sk}^\prime$ with equal linking keys (that is, with $\textbf{sk}_0 = \textbf{sk}^\prime_0$) and with corresponding public keys $\textbf{pk}$ and $\textbf{pk}^\prime$ respectively, any appropriately-sized pair of matrices $\underline{\textbf{pk}}$ and $\underline{\textbf{pk}}^\prime$, if $(\sigma, \mathfrak{T}) \leftarrow \texttt{Sign}(m, \underline{\textbf{pk}}, \textbf{sk})$ and $(\sigma^\prime, \mathfrak{T}^\prime) \leftarrow \texttt{Sign}(m^\prime, \underline{\textbf{pk}}^\prime, \textbf{sk}^\prime)$ and there exists a pair of indices $\ell$ and $\ell^\prime$ such that $\underline{\textbf{pk}}_\ell = \textbf{pk}$ and $\underline{\textbf{pk}}_{\ell^\prime} = \textbf{pk}^\prime$, then
$\texttt{Link}((\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime))=1$.
\end{description}

Note that linkability is only considered with respect to a single component of a matrix entry. When applied to Monero, this is important since only one such entry corresponds to an output public key for double-spend detection purposes.

%In the sequel, we omit $1^\lambda$ and $\textit{par}$ from the input of algorithms for clarity unless there is risk of confusion.

\subsection{Examples}

\begin{example}\label{ex:lsag}
The LSAG signature scheme from \cite{liu2004linkable} is a $1$-LRMS. In this example, $\texttt{Setup}$ always deterministically sets $d:=1$ (so we only use the linking key and there are no auxiliary keys). $\texttt{Setup}$ selects some $G \in \mathbb{G}$ to be a group generator for the group parameters $(p, \mathbb{G}, 1, G)$, two cryptographic hash functions $\mathcal{H}^s:\left\{0,1\right\}^* \to \mathbb{Z}/p\mathbb{Z}$ and $\mathcal{H}^p:\left\{0,1\right\}^* \to \mathbb{G}$. $\texttt{Setup}$ outputs $\textit{par} = (p, \mathbb{G}, 1, G, \mathcal{H}^s, \mathcal{H}^p)$.

$\texttt{KeyGen}$ produces as output a private key $x \in \mathbb{Z}/p\mathbb{Z}$ and a public key $X = xG$. $\texttt{Sign}$ takes as input a private key $y \in \mathbb{Z}/p\mathbb{Z}$, a message $m$, and a ring $\textbf{X} = \left\{X_1, \ldots, X_n\right\}$, and produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature-tag pair $(\sigma, \mathfrak{T})$.

The signature scheme originally described in \cite{liu2004linkable} signs a message $m$ with a ring of keys $\textbf{X} = \left\{X_1, \ldots, X_n\right\}$ and a secret key $x$ corresponding to some $X_\ell$, using the linkability tag (key image) $\mathfrak{T} := x \mathcal{H}^p(\textbf{X})$. Unfortunately, this key image is unsuitable for use in Monero, as changing ring members will change the key image. This would allow the same key to be used twice in two different ring signatures. For use in Monero, we modify this key image to be independent of the non-signing ring members, $\mathfrak{T} := x \mathcal{H}^p(X_\ell)$. This allows these key images to be used for double-spend protection.

An LSAG signature on a message $m$ with ring $\textbf{X}$ using these key images is computed in the following way. First, the signer samples $\alpha, s_{\ell+1}, s_{\ell+2}, \ldots, s_{\ell-1} \in \mathbb{Z}/p\mathbb{Z}$ at random. Next, the signer computes basepoints $H_i = \mathcal{H}^p(X_i)$. Next, the signer computes $c_{\ell+1}  = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid \alpha G \mid \mid \alpha H_\ell)$ and the signer computes each $c_{i+1} = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid s_i G + c_i X_i \mid \mid s_i H_i + c_i \mathfrak{T})$ for $i = \ell+1, \ldots, \ell-1$, naturally identifying index $1$ with index $n+1$. The signer finishes by computing $s_\ell = \alpha - c_\ell x_\ell$ and publishing the signature-tag pair $(\sigma, \mathfrak{T})$ where $\sigma = (c_1, s_1, \ldots, s_n)$.

A purported LSAG signature-tag pair on a message $m$ with ring $\textbf{X}$ is verified in the following way. The verifier sets $c_1^\prime = c_1$ and computes $c_{i+1}^\prime = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid s_i G + c_i^\prime X_i \mid \mid s_i H_i + c_i^\prime \mathfrak{T})$ for $i=1, 2, \ldots, n$. The verifier outputs $1$ when $c_{n+1}^\prime = c_1$ and $0$ otherwise. Note that all verifiers must list the keys in $\textbf{X}$ in an agreed-upon order for the above verification to work; either they should agree upon lexicographic or some other ordering.  LSAG signature-tag pairs are linked merely by comparing key images: two valid signatures with the same key image were signed with the same secret key (and, in the case of Monero, would signal an attempt to double-spend funds).
\end{example}

\begin{remark}
The key image modification in Example \ref{ex:lsag} is due to the basepoint of the key image $\mathfrak{T}$. As noted in \cite{liu2004linkable}, easy variations on key image formats are available. How or whether the security properties of LSAG signatures are retained in practical use given more flexible key image formats, while interesting, is beyond the scope of this work.
\end{remark}

>>>>

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{definition}{Definition}[section]

\title{Compact linkable ring signatures and applications}
\authors{Brandon Goodell\footnote{\texttt{surae.noether@protonmail.com}} and Sarang Noether\footnote{\texttt{sarang.noether@protonmail.com}} and Arthur Blue\footnote{\texttt{randomrun@protonmail.com}}}
\affiliations{Monero Research Lab, independent researcher}
\date{\today}
\type{RESEARCH BULLETIN}
\ident{MRL-0011}

@@

\begin{abstract}
We introduce an efficient linkable ring multisignature construction, \textit{compact linkable spontaneous anonymous group} (CLSAG) signatures. These admit $d$-dimensional keys with a specified linking coordinate but do not have signature sizes directly proportional to $d$. Compared to existing constructions used for signer-ambiguous confidential transactions without trusted setup, CLSAG signatures are smaller and more efficient in terms of both proving and verification time. CLSAG signatures also satisfy some rigorous security definitions: unforgeability depends upon the $k$-OMDL hardness assumption, linkability depends on unforgeability as well as the collision resistance of key aggregation, and signer-ambiguity depends on the DDH assumption. We demonstrate an application for CLSAG signatures for use in transacting multiple assets over ring signature-based transaction protocols.
\end{abstract}

\section{Introduction}
First introduced in \cite{rivest} in the RSA setting and in \cite{liu} in the discrete logarithm setting, ring signatures permit a non-interactive signature on behalf of a set of public keys rather than a single public key. Ring signatures see myriad applications ranging from lightweight anonymous authentication as in \cite{yang2015lightweight} to transaction protocols like Monero in \cite{noether} and CryptoNote in \cite{van2013cryptonote}. A verifier is assured that the signer knows the private key of at least one of these public keys, which are called ring members. Ring signatures are \textit{signer-ambiguous} by nature because the verifier does not learn additional information from the signature about which key is the signer. We stress that methods of practical analysis such as those of  \cite{moser2018empirical,quesnelle2017linkability} can exploit metadata in real-life applications of signer-ambiguous protocols to reduce ambiguity.

Group signature constructions preceding \cite{rivest,liu} require some degree of interactivity, a fixed set of participants, a trusted group manager (or some other trusted setup), or hardness assumptions not based on the discrete logarithm problem. Since \cite{rivest}, ring signatures have enjoyed many improvements, expansions, and modifications. An incomplete list of examples includes: ring signatures are constructed in the bilinear pairing setting in \cite{zhang2002id}, key structures are generalized in \cite{abe20021}, security definitions are improved in \cite{bender}, signature size is improved in \cite{fujisaki2011sub,gu2018constant}, and traceability is introduced in \cite{fujisaki2007traceable}.

Linkable ring signatures were first introduced in \cite{liu}; in the context of distributed ledgers like Monero, linkable ring signatures are the basis for signer-ambiguous transaction authentication.  Linkable ring signatures guarantee that two signatures with the same ring on arbitrary messages can be publicly linked if signed using the same key. An implementation is presented in \cite{liu} in the discrete logarithm setting; that implementation functions for similar reasons as Schnorr signatures in \cite{schnorr1991efficient}.

The key images in \cite{liu} are unsuitable for applications where signatures must be linked key-by-key, not ring-by-ring (such as for ``double-signing'' protection in a setting where users generate new keys over time and select \textit{ad hoc} ring members). Resistance to double-spend attempts is ensured using key images as described in \cite{van2013cryptonote}. More recent work of \cite{noether} extends the approach of \cite{liu} to enable a signer-ambiguous confidential transaction model with \textit{ad hoc} ring member selection. In \cite{noether}, transaction amounts are replaced with Pedersen commitments to amounts together with range proofs. Signatures are constructed from key vectors including differences of amount commitments as one of the keys. However, the proofs in \cite{noether} are informal and not based on rigorous security models.

Alternatives to ring signatures like more general zero-knowledge proving systems typically require a trusted party to honestly perform a setup process (as in \cite{groth,ben2014succinct,groth2018updatable}) or lack practical efficiency for large circuits (as in \cite{bulletproofs}), meaning that such systems may not be appropriate for distributed ledger applications. However, more recent approaches such as \cite{bowehalo,hoffmann2019efficient} show both improvements to the trust requirement as well as improvements in efficiency.

\subsection{Our contribution}
We first introduce a formal definition of $d$-dimensional linkable ring signatures. We present an implementation of a new signature scheme, $d$-CLSAG signatures. These have $d$-dimensional keys and are \textit{compact} linkable spontaneous anonymous group signatures in the sense that signature size scales with the \textit{sum} of ring size and the dimension $d$. Equivalent MLSAG signatures of \cite{noether} produce signatures that scale with the product of ring size and $d$ rather than the sum. Size efficiency comes from an aggregation of keys across components, similar to the approach from \cite{maxwell2018simple,qian2010non}, resulting in $d$-CLSAG signatures that are about half the size of \cite{noether}, and can be generated and verified more quickly.

We present linkable ring signatures and some examples in Section \ref{sec:lrs-and-exs}. We present a compact $d$-LRS scheme we call $d$-CLSAG in Section \ref{sec:themainthing} and measure efficiency. We present an informal description of a transaction protocol for confidentially transacting multiple asset types simultaneously, which we call a \textit{multi-asset ring confidential transaction} or MARCT in Section \ref{sec:marct}. We reserve discussion of security for the appendix, where we make a new definition for unforgeability in linkable ring signatures that takes into account both insider corruption and forgeries from partially-corrupted rings. We prove that $d$-CLSAG signatures are unforgeable up to the hardness of the $k$-one-more discrete logarithm ($k$-OMDL) problem under this definition in the random oracle model in Appendix \ref{app:unf}.  We also make some comments on signer ambiguity and linkability in Appendix \ref{app:link-and-sa}.

\subsection{Notation}\label{sec:notation}
We denote algorithms with typefont majuscule English letters like $\texttt{A}$, $\texttt{B}$, or $\texttt{O}$, or typefont names like $\texttt{Setup}$, $\texttt{KeyGen}$, and so on.  For any prime $p$, denote the field with $p$ elements as $\mathbb{F}_p := \mathbb{Z}/p\mathbb{Z}$, and denote the non-zero elements as $\mathbb{F}_p^*$.

Group parameters are denoted as a tuple $(p, \mathbb{G}, d, G)$ where $\mathbb{G}$ is an elliptic curve group with prime order $p$, $d$ is a dimension, and $G$ is a generator of $\mathbb{G}$. We denote integers, bits, indices, and scalars in $\mathbb{F}_p$ with minuscule English letters $x, y, z, b, c, i, j, k$, etc.\ and we denote group elements with majuscule English letters, $G, X, W$, and so on. We use miniscule Greek letters like $\sigma$ to describe signatures and majuscule calligraphic Latin letters like $\mathfrak{T}$ when describing key images.

For these group parameters, the secret key space is $\mathbb{F}_p^*$ and the public key space is $\mathbb{G}$. For any non-zero secret key $\textit{sk} \in \mathbb{F}_p^*$, the corresponding public key $\textit{pk}$ is computed from the generator $G$ through exponentiation in $\mathbb{G}$ as usual. However, we use notation for a module over the field $\mathbb{Z}_p$ to maintain consistency with, say \cite{van2013cryptonote} and \cite{noether}.

@@

We distinguish oracles with calligraphic font, e.g.\ $\mathcal{CO}$ denotes a corruption oracle, $\mathcal{SO}$ denotes a signing oracle. If the codomain of a random oracle is the field of scalars $\mathbb{F}_p$, we denote this $\mathcal{H}^s$ (hash-to-scalar). If the codomain is $\mathbb{G}$, we denote this $\mathcal{H}^p$ (hash-to-point).

\section{Linkable ring multisignatures}\label{sec:lrs-and-exs}
In this section, we recall linkable ring signature (LRS) schemes, definitions of correctness in verification and linkability, and we provide two examples.

\subsection{LRS schemes}
\begin{definition}[LRS]\label{def:lrs}
A \textit{linkable ring signature scheme} is a tuple $\left(\texttt{Setup}, \texttt{KeyGen}, \right. \allowbreak \left. \texttt{Sign}, \texttt{Verify}, \texttt{Link}\right)$ satisfying the following. 
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ takes as input a security parameter $1^\lambda$, produces some public parameters $\textit{par}$.

@@

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, sk)\right) \to \left\{\bot_{\texttt{Sign}}, \sigma \right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, an arbitrary message $m \in \left\{0,1\right\}^*$, an ad hoc \textit{ring} of public keys $\textbf{pk} = \left\{pk_1, \ldots, pk_n\right\}$, and a secret key $sk$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\textit{out} = \bot_{\texttt{Sign}}$ or a signature $\textit{out} = \sigma$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, \sigma)\right) \to \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a ring of public keys $\textbf{pk}$, and a signature $\sigma$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$; $0$ indicates the signature is not verified, and $1$ indicates the signature is verified.

\item $\texttt{Link}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, \sigma), (m^\prime, \textbf{pk}^\prime, \sigma^\prime)\right)$. $\texttt{Link}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, and a pair of tuples $(m, \textbf{pk}, \sigma), (m^\prime, \textbf{pk}^\prime, \sigma^\prime)$ for messages $m, m^\prime$, rings $\textbf{pk}$, $\textbf{pk}^\prime$, and ring signatures $\sigma, \sigma^\prime$. $\texttt{Link}$ produces as output a bit $b \in \left\{0,1\right\}$; $0$ indicates the signatures are not linked or invalid, and $1$ indicates the signatures are valid and linked.\footnote{This is the opposite of the conventions in, say, \cite{tsang2004separable}, which outputs $0$ to indicate two signatures are linked (\textit{i.e.}\ rejected) and outputs $1$ to indicate two signatures are not linked (\textit{i.e.}\ accepted).} 
\end{itemize}
\end{definition}

In the sequel, we implicitly assume all algorithms in a LRS take $1^\lambda$ as input, and all algorithms (except $\texttt{Setup}$) takes $\textit{par}$ as input. We suppress this notation in the sequel. For example, we write $\textit{par} \leftarrow \texttt{Setup}$ instead of $\textit{par} \leftarrow \texttt{Setup}(1^\lambda)$ and $\sigma \leftarrow \texttt{Sign}(m, \textbf{pk}, \textit{sk})$ instead of $\sigma \leftarrow \texttt{Sign}(1^\lambda, \textit{par}, m, \textbf{pk}, \textit{sk})$.

Note the dimension of keys in Definition \ref{def:lrs} is not specified. If the keys from $\texttt{KeyGen}$ have dimension $d > 1$, we instead say the LRS is a $d$-LRS and use the vector notation introduced in Section \ref{sec:notation} representing keys in boldface (\textit{e.g.}\ \textbf{pk} instead of \textit{pk}) and rings in underlined boldface ($\underline{\textbf{pk}}$ instead of $\textbf{pk}$). We always assume the first coordinate of a secret key is the \textit{linking coordinate}, and linkability depends only on the linking coordinate of the signing key vector. For ring confidential transactions, this is important since only one such entry corresponds to an output public key for double-spend detection purposes and the rest of the keys could have been adversarially generated.

\begin{definition}[Correctly verified]
Let $m$ be any message, $\textit{sk}$ be any secret key with corresponding public key $\textit{pk}$, and let $\textbf{pk}$ be a multiset of public keys $\textbf{pk} = \left\{\textit{pk}_1, \ldots, \textit{pk}_n\right\}$. If there exists an index $1 \leq \ell \leq n$ satisfying $\textit{pk} = \textit{pk}_\ell$, then \[\texttt{Verify}(m, \textbf{pk}, \texttt{Sign}(m, \textbf{pk}, \textit{sk})) = 1.\]
\end{definition}
Soundness in verification is the property of unforgeability. See Appendix \ref{app:unf}.

\begin{definition}[Correctly linkable]
Let $m, m^\prime$ be any messages, $\textit{sk}, \textit{sk}^\prime$ be any secret keys with corresponding public keys $\textit{pk}$, $\textit{pk}^\prime$. Let $\textbf{pk}$, $\textbf{pk}^\prime$ be multisets of public keys with indices $\ell, \ell^\prime$, respectively, such that $\textbf{pk}_\ell = \textit{pk}$ and $\textbf{pk}^\prime_{\ell^\prime} = \textit{pk}^\prime$. Let $\textit{sk}, \textit{sk}^\prime$ be the respective private keys for $\textit{pk}, \textit{pk}^\prime$ with respective linking coordinates $\textit{sk}_0$ and $\textit{sk}^\prime_0$. Correct linkability means being both \textit{positively} and \textit{negatively} linkable, in the following sense.
\begin{itemize}
\item \textbf{Positively linkable:} If the linking coordinates satisfy $\textit{sk}_0 = \textit{sk}^\prime_0$, then \[\textit{Link}((m, \textbf{pk}, \texttt{Sign}(m, \textbf{pk}, sk)), (m^\prime, \textbf{pk}^\prime, \texttt{Sign}(m^\prime, \textbf{pk}^\prime, sk^\prime)))=1\] except with negligible probability.

\item \textbf{Negatively linkable:} If the linking coordinates satisfy $\textit{sk}_0 \neq \textit{sk}^\prime_0$, then \[\textit{Link}((m, \textbf{pk}, \texttt{Sign}(m, \textbf{pk}, sk)), (m^\prime, \textbf{pk}^\prime, \texttt{Sign}(m^\prime, \textbf{pk}^\prime, sk^\prime)))=0\] except with negligible probability.
\end{itemize}
\end{definition}

Soundness in linkability means informally that an adversarial algorithm can only produce tuples $y=(m, \textbf{pk}, \sigma)$ and $y^\prime =(m^\prime, \textbf{pk}^\prime, \sigma^\prime)$ such that $\texttt{Link}(y, y^\prime)=1$ by computing $\sigma$, $\sigma^\prime$ from $\texttt{Sign}$ using secret keys $\textit{sk}, \textit{sk}^\prime$ with matching linking coordinates. See Appendix \ref{app:link-and-sa}.

\subsection{Examples}
In the following examples, all verifiers must list the keys in $\textbf{pk}$ in an agreed-upon order for the above verification to work; either they should agree upon lexicographic or some other ordering.  Linking occurs merely by comparing key images: two valid signatures with the same key image were signed with the same secret key (and, in transaction applications, would signal an attempt to double-spend funds).

\begin{example}\label{ex:lsag}
The signature scheme of \cite{liu} is an LRS; we present the key image variant from \cite{van2013cryptonote} with a key image appropriate for linking signatures key-by-key.  The signature scheme originally described in \cite{liu} signs a message $m$ with a ring of keys $\textbf{pk} = \left\{\textit{pk}_1, \ldots, \textit{pk}_n\right\}$ and a secret index-key pair $(\ell, \textit{sk})$ corresponding to some $\textit{pk}_\ell$, using the key image $\mathfrak{T} := \textit{sk} \cdot \mathcal{H}^p(\textbf{pk})$.

Unfortunately, this key image is unsuitable for transaction applications, as changing ring members will change the key image, allowing the same key to sign twice. For use in a transaction protocol and following \cite{van2013cryptonote}, we modify this key image from that of \cite{liu} to be $\mathfrak{T} := \textit{sk} \cdot  \mathcal{H}^p(\textit{pk}_\ell)$, which is independent of the non-signing ring members. This allows these key images to be used for double-spend protection, as discussed previously.

$\texttt{Setup}$ always deterministically sets $d:=1$ so we only use the linking key and there are no auxiliary keys. $\texttt{Setup}$ selects a generator $G \in \mathbb{G}$ to be a group generator for the group parameters $(p, \mathbb{G}, d, G)$, two cryptographic hash functions $\mathcal{H}^s:\left\{0,1\right\}^* \to \mathbb{F}_p$ and $\mathcal{H}^p:\left\{0,1\right\}^* \to \mathbb{G}$. $\texttt{Setup}$ outputs $\textit{par} = (p, \mathbb{G}, d, G, \mathcal{H}^s, \mathcal{H}^p)$.

$\texttt{KeyGen}$ produces as output a secret key $\textit{sk} \in \mathbb{F}_p^*$ and the corresponding public key $\texttt{pk} = \textit{sk}\cdot G\in \mathbb{G}$.

$\texttt{Sign}$ takes as input a (non-zero) private key $\textit{sk} \in \mathbb{F}_p^*$, a message $m$, a ring $\textbf{pk} = \left\{\textit{pk}_1, \ldots, \textit{pk}_n\right\}$, and produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature $\sigma$, computed as follows. First, the signer samples $\alpha, s_{\ell+1}, s_{\ell+2}, \ldots, s_{\ell-1} \in \mathbb{F}_p$ at random. Next, the signer computes basepoints $H_i = \mathcal{H}^p(\textit{pk}_i)$ and the key image $\mathfrak{T} = \textit{sk} \cdot \mathcal{H}^p(\textit{pk}_\ell)$, the first challenge $$c_{\ell+1}  = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid \alpha G \mid \mid \alpha H_\ell)$$ and each subsequent challenge $$c_{i+1} = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid s_i G + c_i \cdot \textit{pk}_i \mid \mid s_i H_i + c_i \mathfrak{T})$$ for $i = \ell+1, \ldots, \ell-1$, naturally identifying index $1$ with index $n+1$. The signer finishes by computing $s_\ell = \alpha - c_\ell \cdot \textit{sk}$ and publishing the signature $\sigma$ where $\sigma = (c_1, s_1, \ldots, s_n, \mathfrak{T})$.

$\texttt{Verify}$ takes as input a message $m$, a ring $\textbf{pk}$, and a purported signature $\sigma^\prime$. $\texttt{Verify}$ parses the signature $(c_1, s_1, \ldots, s_n, \mathfrak{T}) \leftarrow \sigma$. If $c_1 \notin \mathbb{F}_p$ or any $s_i \notin \mathbb{F}_p$ or $\mathfrak{T} \notin \mathbb{G}$, then $\texttt{Verify}$ outputs $0$. Otherwise, the verifier computes $H_i := \mathcal{H}^p(X_i)$ for each ring member, sets $c_1^\prime := c_1$, and computes the challenges $$c_{i+1}^\prime = \mathcal{H}^s(\textbf{pk} \mid \mid m \mid \mid s_i G + c_i^\prime \cdot \textit{pk}_i \mid \mid s_i H_i + c_i^\prime \mathfrak{T})$$ for $i=1, 2, \ldots, n$. The verifier outputs $1$ when $c_{n+1}^\prime = c_1$ and $0$ otherwise.

$\texttt{Link}$ checks the validity of both signatures. If both are valid, $\texttt{Link}$ parses the key images $\mathfrak{T}$ and $\mathfrak{T}^\prime$. If either are not in $\mathbb{G}$, the linker outputs $0$ and terminates. Otherwise, the linker outputs $1$ when $\mathfrak{T} = \mathfrak{T}^\prime$ and $0$ otherwise.
\end{example}

\begin{remark}
The key image modification in Example \ref{ex:lsag} is due to the basepoint of the key image $\mathfrak{T}$. As noted in \cite{liu}, variations on key image formats may be desirable. How or whether the security properties of LSAG signatures are retained in practical use given more flexible key image formats, while interesting, is beyond the scope of this work.
\end{remark}
